---
title: "Colorimentric output analysis"
author: "Andrea Clausen Lind"
date: "2023/05/04"
output: html_notebook
---

## Load libraries
```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(viridis)
library(hrbrthemes)
library(broom)
library(patchwork)

filter <- dplyr::filter
```

## Load data files
Load data. Adapted to the data format output by the Mars analysis program.(RUN ONCE)
```{r,eval=FALSE}
#Extract file names and paths
Dir_path <- "./2023-06-01_Absorbance_Fluo/"
File_path_list <- list.files(Dir_path, full.names = T)
File_list <- list.files(Dir_path)

sample_time <- 
  File_list %>% 
  as_tibble() %>% 
  mutate(value = gsub("h.xlsx","",value)) %>% 
  mutate(value = gsub(".*Abs_","",value)) %>% 
  rename("time" = "value") %>% 
  pull(time)

tbl_full <- tibble()

# Loop through and load each file one by one
for (i in 1:length(File_list)) {
  
  FileToRead <- File_path_list[i]
  FileName <- File_list[i]
  df_file <- read_xlsx(FileToRead, skip = 11, col_names = FALSE) 
  df_file <- 
    df_file %>%  
    select(-"...1", -"...2") %>%
    distinct() %>%
    t() %>%
    row_to_names(row_number = 1) %>%
    as_tibble() %>%
    select(-"Blank B") %>%
    rename(wavelength_nm = "Wavelength [nm]") %>% 
    pivot_longer(!wavelength_nm, names_to = "sample",values_to = "absorbance") %>% 
    mutate(sample = gsub(" ", "_",sample), sample = factor(sample))
  
  df_file <- 
    df_file %>% 
    mutate(time = sample_time[i])
  
  tbl_full <- bind_rows(tbl_full,df_file)
}


# Add metadata
df_meta =
  read_csv("Metadata.csv") %>% 
  mutate(sample = gsub(" ","_",sample))

tbl_complete <- 
  tbl_full %>% 
  full_join(df_meta,by = "sample")

# Tweak columns 
tbl_complete <- 
  tbl_complete %>% 
  mutate(time = as.integer(time)) %>% 
  mutate(absorbance = as.numeric(absorbance)) %>% 
  mutate(induction = factor(induction,levels = c("0 uM","1 uM"), ordered = T)) %>% 
  mutate(wavelength_nm = as.double(as.character(wavelength_nm))) %>% 
  mutate(strain = factor(strain,levels = c("AMG0-1","AlCG0-1","APDG0-1","APVG0-1"),ordered = T))

head(tbl_complete)

saveRDS(tbl_complete,file = "./rds_data/tbl_complete.rds")

```

Load processed data instead
```{r}
tbl_complete <- readRDS("./rds_data/tbl_complete.rds")
```


Add unique ID to each biological replicate
```{r}
tbl_complete <- 
  tbl_complete %>% 
  mutate(sample_ID = paste(strain,time,induction,replicate,sep = "_")) %>% 
  relocate(sample_ID,.before = wavelength_nm)
```

Extract sensitive and robust wavelengths. Adjust absorbance based on OD correction function. POSTPONE DUE TO LARGE VARIANCE CAUSED BY SMALL DIFFERNECES IN ABSORBANCE
```{r}
# Postpone OD correction adjustment.
 tbl_data <-
   tbl_complete %>% 
   filter(wavelength_nm %in% c("395", "450", "520", "600", "750")) %>% 
   mutate(abs_adj = case_when(
     wavelength_nm == "395" ~ (0.4832469*absorbance)/(2.5931168-absorbance),
     wavelength_nm == "450" ~ (0.4645766*absorbance)/(2.4481330-absorbance),
     wavelength_nm == "520" ~ (0.4541800*absorbance)/(2.3384806-absorbance),
     wavelength_nm == "600" ~ (0.4587302*absorbance)/(2.2147928-absorbance),
     wavelength_nm == "750" ~ (0.4594981*absorbance)/(2.0908275-absorbance))) 


#Plot both asjusted and unadjusted data to compare and confirm that everyting looks as expected
plot_adjust_data <-
  tbl_data %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate)) %>%
  mutate(wavelength_nm = paste0(wavelength_nm," nm")) %>% 
  ggplot(aes(
    x = time,
    y = abs_adj,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
  facet_grid(vars(wavelength_nm), vars(strain)) +
   theme_bw() +
   theme(aspect.ratio = 1) +
  ylim(0,3) +
  coord_fixed() +
  ylab("Absorbance [a.u.]") +
  xlab("Time [h]") +
  ggtitle("Data adjusted for absorbance saturation")

plot_adjust_data

plot_unadjust_data <-
tbl_data %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate)) %>%
  mutate(wavelength_nm = paste0(wavelength_nm," nm")) %>% 
  ggplot(aes(
    x = time,
    y = absorbance,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
  facet_grid(vars(wavelength_nm), vars(strain)) +
   theme_bw() +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylim(0,3) +
  ylab("Adjusted absorbance [a.u.]") +
  xlab("Time [h]") +
  ggtitle("Unadjusted data")

plot_unadjust_data
```

```{r, eval = FALSE}
plot_abs_adjust <-
  plot_unadjust_data + plot_adjust_data

ggsave(plot_abs_adjust, filename = "Figures/absAdjust.pdf",width = 15)
```

Calculate absorbance by colorimetric compounds and plot Note that inflation of certain data ponts using the OD saturation correction resulted in these being out of range of the dataset - in some cases over 1000 - showing that with data this noisy this method for correction of saturated ODs cannot be applied, as small differences caused by pipetting or measurement errors caused by the microplate analysis being taken for one well at the time. Cannot use this method for correction. Will use relative abundance of the compounds instead, compared to a negative control. 
```{r}
tbl_data_lyc <-
  tbl_data %>% 
  filter(strain %in% c("AMG0-1", "AlCG0-1"),
         wavelength_nm %in% c("395","520", "600")) %>% 
  select(-abs_adj) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = absorbance) %>%
  mutate(A = 395*((1-(10^(abs395nm-abs600nm)))/((395/600)-10^(abs395nm-abs600nm)))) %>% 
  mutate(B = abs395nm+log10(1-(A/600))) %>% 
  mutate(ODscat = B-log10(1-(A/520))) %>% 
  mutate(AbsLyc = abs520nm-ODscat) 

tbl_data_lyc_adj <-
  tbl_data %>% 
  filter(strain %in% c("AMG0-1", "AlCG0-1"),
         wavelength_nm %in% c("395","520", "600")) %>%
  select(-absorbance) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = abs_adj) %>%
  mutate(A = 395*((1-(10^(abs395nm-abs600nm)))/((395/600)-10^(abs395nm-abs600nm)))) %>% 
  mutate(B = abs395nm+log10(1-(A/600))) %>% 
  mutate(ODscat = B-log10(1-(A/520))) %>% 
  mutate(AbsLyc = abs520nm-ODscat) 

tbl_data_lyc %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>% 
  ggplot(aes(
    x = time,
    y = AbsLyc,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell (LPC) unadjusted") +
  xlab("Time [h]") +
  ggtitle("Unadjusted lycopene data")

tbl_data_lyc_adj %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>% 
  ggplot(aes(
    x = time,
    y = AbsLyc,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell (LPC) adjusted") +
  xlab("Time [h]") +
  ggtitle("Adjusted lycopene data")

```
The non-normalized data does show differences between the induced and uninduced strains for AlCG0-1 strating at 5h. However, the OD saturation makes it impossible to compare the relative Lycopene amount between the strains. The OD correction is however extremely sensitive to variations in the data and inflates measurements points based on small differences in OD, see below. Will try to readjust the OD correction function, but it's unlikely to help. An alternative method could also be to calculate the relative lycopene amount based on the unadjusted data, by subtracting the mean absorbance of the output null strain from that of the lycopene strain. 
```{r}
#Calculate means and standard deviations for lycopene
tbl_data_lyc_mean <-
  tbl_data_lyc %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsLyc = mean(AbsLyc),
         sd_AbsLyc = sd(AbsLyc)) %>% 
  ungroup() 

tbl_data_lyc_adj_mean <-
  tbl_data_lyc_adj %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsLyc = mean(AbsLyc),
         sd_AbsLyc = sd(AbsLyc)) %>% 
  ungroup() 

tbl_data_lyc_mean %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>% 
  ggplot(aes(
    x = time,
    y = mean_AbsLyc,
    ymin = mean_AbsLyc-sd_AbsLyc,
    ymax = mean_AbsLyc+sd_AbsLyc,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_ribbon(alpha =0.3) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell") +
  xlab("Time") +
  ggtitle("Unadjusted data")

tbl_data_lyc_adj_mean %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>%
  ggplot(aes(
    x = time,
    y = mean_AbsLyc,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell") +
  xlab("Time") +
  ggtitle("Adjusted data")

```

Caluclate difference between lycopene strain and null strain in induced/uninduced condition at different timepoints
```{r}
tbl_lyc_diff <-
  tbl_data_lyc_mean %>% 
  select(-c(abs395nm,abs520nm,abs600nm, A, B, ODscat, sample, sample_ID)) %>% 
  distinct() %>%  
  group_by(time, induction,replicate) %>% 
  mutate(diff_AbsLyc = AbsLyc-AbsLyc[strain=="AMG0-1"]) %>% 
  ungroup() %>%   
  group_by(time, induction) %>% 
  mutate(diff_mean_AbsLyc = mean_AbsLyc-mean_AbsLyc[strain=="AMG0-1"],
         diff_sd_AbsLyc = sqrt((sd_AbsLyc)^2+(sd_AbsLyc[strain=="AMG0-1"])^2)) %>% 
  ungroup() 

tbl_lyc_diff %>% 
  filter(strain == "AlCG0-1") %>% 
  mutate(group_rep = paste0(induction,"_",strain)) %>%
  ggplot(aes(
    x = time,
    y = diff_mean_AbsLyc,
    ymin = diff_mean_AbsLyc-diff_sd_AbsLyc,
    ymax = diff_mean_AbsLyc+diff_sd_AbsLyc,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_ribbon(aes(), alpha=0.3) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("lycopene per cell compared negative control") +
  xlab("Time") +
  ggtitle("Unadjusted data")

plot_diff_lyc <-
  tbl_lyc_diff %>%
  filter(strain == "AlCG0-1") %>%
  mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
      x = time,
      y = diff_mean_AbsLyc,
      ymin = diff_mean_AbsLyc - diff_sd_AbsLyc,
      ymax = diff_mean_AbsLyc + diff_sd_AbsLyc
    )
  ) +
  geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsLyc, 
        group = induction),
    size = 0.5,
    color = "black",
    alpha = 0.4,
    position = position_dodge(width = 0.9)
  ) +
    scale_fill_manual(values = c("grey", "darkorange"), name = "Induction") +
#geom_errorbar(aes(fill = induction),width=0.2, position=position_dodge(0.9)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" lycopene absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean AlCG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_lyc
```

```{r,eval=F}
ggsave(plot_diff_lyc, filename = "Figures/deltaLyc.pdf",width = 7)
```

Proceed with data from prodeoxy- and proviolacein strains. These can be analysed together as sensitive and robust wavelengths are the same.
Tried to calculate prodeoxy-/provioalacein per cell using the adjusted ODs, but again the noise in the data causes infalition of certain measurement points. Calculate the difference compared to the null strain instead, using the mean. 
```{r}
tbl_data_pro <-
  tbl_data %>% 
  filter(strain %in% c("AMG0-1", "APDG0-1","APVG0-1"),
         wavelength_nm %in% c("395","450", "750")) %>% 
  select(-abs_adj) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = absorbance) %>%
  mutate(A = 395*((1-(10^(abs395nm-abs750nm)))/((395/750)-10^(abs395nm-abs750nm)))) %>% 
  mutate(B = abs395nm+log10(1-(A/750))) %>% 
  mutate(ODscat = B-log10(1-(A/450))) %>% 
  mutate(AbsPro = abs450nm-ODscat) 

tbl_data_pro_adj <-
  tbl_data %>% 
  filter(strain %in% c("AMG0-1", "APDG0-1","APVG0-1"),
         wavelength_nm %in% c("395","450", "750")) %>% 
  select(-absorbance) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = abs_adj) %>%
  mutate(A = 395*((1-(10^(abs395nm-abs750nm)))/((395/750)-10^(abs395nm-abs750nm)))) %>% 
  mutate(B = abs395nm+log10(1-(A/750))) %>% 
  mutate(ODscat = B-log10(1-(A/450))) %>% 
  mutate(AbsPro = abs450nm-ODscat) 

tbl_data_pro %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>% 
  ggplot(aes(
    x = time,
    y = AbsPro,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell (LPC) unadjusted") +
  xlab("Time [h]") +
  ggtitle("Unadjusted lycopene data")

tbl_data_pro_adj %>% 
  mutate(group_rep = paste0(induction, "_" ,replicate,"_",strain)) %>% 
  ggplot(aes(
    x = time,
    y = AbsPro,
    color = induction
  )) +
  geom_line(aes(group = group_rep),size = 0.5, alpha = 0.7) +
  geom_point(size = 0.55) +
  scale_color_ipsum() +
   theme_bw() +
  facet_wrap(vars(strain)) +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab("Lycopene per cell (LPC) adjusted") +
  xlab("Time [h]") +
  ggtitle("Adjusted lycopene data")


```
Calculate the mean absorbance for each strain.  
Subtract null strain values for unduced/uninduced prodeoxy-/proviolacein strains to gett difference after removal of baseline. Do this for unadjusted data only due excessive noise in data adjusted for OD saturation. 
```{r}
# Calcuclate mean
tbl_data_pro_mean <-
  tbl_data_pro %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsPro = mean(AbsPro),
         sd_AbsPro = sd(AbsPro)) %>% 
  ungroup() 

tbl_data_pro_adj_mean <-
  tbl_data_pro_adj %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsPro = mean(AbsPro),
         sd_AbsPro = sd(AbsPro)) %>% 
  ungroup() 

#Plot mean values with standard deviations for unadjusted data
plot_mean_pro <-
  tbl_data_pro_mean %>% 
  mutate(group_rep = paste0(induction,"_",strain)) %>% 
  mutate(time = as.factor(time)) %>% 
  ggplot(aes(
    x = time,
    y = mean_AbsPro,
    ymin = mean_AbsPro-sd_AbsPro,
    ymax = mean_AbsPro+sd_AbsPro, 
    color = induction)) +
  geom_line(aes(group=group_rep),size = 0.5, alpha = 0.7) +
  geom_ribbon(aes(color = induction), alpha =0.3) +    
  geom_errorbar(aes(fill = induction),width=0.2) +
  geom_point(size = 0.55) +
  facet_wrap(vars(strain)) +
  theme_bw() +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression("Pro(deoxy)violacein absorbance")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean pro(deoxy)violacein absorbance (unadjusted OD)"))

plot_mean_pro

#Plot mean values with standard deviations for adjusted data
plot_mean_pro_adj <-
  tbl_data_pro_adj_mean %>% 
  mutate(group_rep = paste0(induction,"_",strain)) %>% 
  mutate(time = as.factor(time)) %>% 
  ggplot(aes(
    x = time,
    y = mean_AbsPro,
    ymin = mean_AbsPro-sd_AbsPro,
    ymax = mean_AbsPro+sd_AbsPro, 
    color = induction)) +
  geom_line(aes(group=group_rep),size = 0.5, alpha = 0.7) +
  geom_ribbon(aes(color = induction), alpha =0.3) +    
  geom_errorbar(aes(fill = induction),width=0.2) +
  geom_point(size = 0.55) +
  facet_wrap(vars(strain)) +
  theme_bw() +
   theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression("Pro(deoxy)violacein absorbance")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean pro(deoxy)violacein absorbance (adjusted OD)"))

plot_mean_pro_adj

# Calculate and plot difference
tbl_pro_diff <-
  tbl_data_pro_mean %>% 
  select(-c(abs395nm,abs450nm,abs750nm, A, B, ODscat, sample, sample_ID)) %>% 
  distinct() %>%
    group_by(time, induction, replicate) %>% 
  mutate(diff_AbsPro = AbsPro-AbsPro[strain=="AMG0-1"]) %>% 
  ungroup() %>%   
  group_by(time, induction) %>% 
  mutate(diff_mean_AbsPro = mean_AbsPro-mean_AbsPro[strain=="AMG0-1"]) %>% 
  mutate(diff_sd_AbsPro = sqrt((sd_AbsPro)^2+(sd_AbsPro[strain=="AMG0-1"])^2)) %>% 
  ungroup() 

plot_diff_pdeox <-
  tbl_pro_diff %>%
  filter(strain == "APDG0-1") %>%
  mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro)) +
  geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        group = induction),
    size = 0.5,
    color = "black",
    alpha = 0.4,
    position = position_dodge(width = 0.9)
  ) +
  scale_fill_manual(values = c("grey", "darkgreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" prodeoxyviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APDG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pdeox

plot_diff_pvio <-
  tbl_pro_diff %>%
  filter(strain == "APVG0-1") %>%
  mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro)) +
  geom_bar(
    aes(fill = induction),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        group = induction),
    size = 0.5,
    color = "black",
    alpha = 0.4,
    position = position_dodge(width = 0.9)
  ) +
  scale_fill_manual(values = c("grey", "darkolivegreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" proviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APVG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pvio




```



Patch together plots and save
```{r, eval=FALSE}
plot_diff_bar <-
  plot_diff_lyc + plot_diff_pdeox + plot_diff_pvio

ggsave(plot_diff_bar, filename = "Figures/deltaBarAll.pdf",width = 15)
```

```{r}

#Try plotting the diff as lines instead
plot_diff_lyc_line <-
  tbl_lyc_diff %>%
  filter(strain == "AlCG0-1") %>%
  #mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
      x = time,
      y = diff_mean_AbsLyc,
      ymin = diff_mean_AbsLyc - diff_sd_AbsLyc,
      ymax = diff_mean_AbsLyc + diff_sd_AbsLyc,
      color = induction
    )
  ) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsLyc, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_fill_manual(values = c("grey", "darkorange"), name = "Induction") +
  scale_color_manual(values = c("grey", "darkorange"), name = "Induction") +
  #geom_errorbar(aes(fill = induction),width=0.2, position=position_dodge(0.9)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" lycopene absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean AlCG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_lyc_line

plot_diff_pdeox_line <-
  tbl_pro_diff %>%
  filter(strain == "APDG0-1") %>%
 # mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro,
    color = induction)) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_color_manual(values = c("grey", "darkgreen"), name = "Induction") +
  scale_fill_manual(values = c("grey", "darkgreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" prodeoxyviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APDG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pdeox_line


plot_diff_pvio_line <-
  tbl_pro_diff %>%
  filter(strain == "APVG0-1") %>%
 # mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro,
    color = induction)) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_color_manual(values = c("grey", "darkolivegreen"), name = "Induction") +
  scale_fill_manual(values = c("grey", "darkolivegreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" proviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APVG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pvio_line

```
Patch together plots and save
```{r, eval = FALSE}
plot_diff_line <-
  plot_diff_lyc_line + plot_diff_pdeox_line + plot_diff_pvio_line

ggsave(plot_diff_line, filename = "Figures/deltaLineAll.pdf",width = 15)
```

Attempt using alternative robust and sensitive wavelengths for lycopene respectively prodeoxy an provioalcein data, excluding measurements below 450 to reduce noise and trying out 650 nm as a sensitive wavelength for prodeoxy-and proviolacein. 
```{r}
tbl_data2 <-
   tbl_complete %>% 
   filter(wavelength_nm %in% c("550", "450", "520", "600", "700"))

tbl_data_lyc2 <-
  tbl_data2 %>% 
  filter(strain %in% c("AMG0-1", "AlCG0-1"),
         wavelength_nm %in% c("700","520", "600")) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = absorbance) %>%
  mutate(A = 700*((1-(10^(abs700nm-abs600nm)))/((700/600)-10^(abs700nm-abs600nm)))) %>% 
  mutate(B = abs700nm+log10(1-(A/600))) %>% 
  mutate(ODscat = B-log10(1-(A/520))) %>% 
  mutate(AbsLyc = abs520nm-ODscat) 

tbl_data_pro2 <-
  tbl_data2 %>% 
  filter(strain %in% c("AMG0-1", "APDG0-1","APVG0-1"),
         wavelength_nm %in% c("550","450", "700")) %>% 
  mutate(wavelength_nm = paste0("abs",wavelength_nm,"nm")) %>% 
  pivot_wider(names_from = wavelength_nm, values_from = absorbance) %>%
  mutate(A = 550*((1-(10^(abs550nm-abs700nm)))/((550/700)-10^(abs550nm-abs700nm)))) %>% 
  mutate(B = abs550nm+log10(1-(A/700))) %>% 
  mutate(ODscat = B-log10(1-(A/450))) %>% 
  mutate(AbsPro = abs450nm-ODscat) 

tbl_data_lyc_mean2 <-
  tbl_data_lyc2 %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsLyc = mean(AbsLyc),
         sd_AbsLyc = sd(AbsLyc)) %>% 
  ungroup() 

tbl_data_pro_mean2 <-
  tbl_data_pro2 %>% 
  group_by(strain, induction, time) %>% 
  mutate(mean_AbsPro = mean(AbsPro),
         sd_AbsPro = sd(AbsPro)) %>% 
  ungroup() 

tbl_lyc_diff2 <-
  tbl_data_lyc_mean2 %>% 
  select(-c(abs520nm,abs600nm,abs700nm, A, B, ODscat, sample, sample_ID)) %>% 
  distinct() %>%  
  group_by(time, induction,replicate) %>% 
  mutate(diff_AbsLyc = AbsLyc-AbsLyc[strain=="AMG0-1"]) %>% 
  ungroup() %>%   
  group_by(time, induction) %>% 
  mutate(diff_mean_AbsLyc = mean_AbsLyc-mean_AbsLyc[strain=="AMG0-1"],
         diff_sd_AbsLyc = sqrt((sd_AbsLyc)^2+(sd_AbsLyc[strain=="AMG0-1"])^2)) %>% 
  ungroup() 

tbl_pro_diff2 <-
  tbl_data_pro_mean2 %>% 
  select(-c(abs550nm,abs450nm,abs700nm, A, B, ODscat, sample, sample_ID)) %>% 
  distinct() %>%
    group_by(time, induction, replicate) %>% 
  mutate(diff_AbsPro = AbsPro-AbsPro[strain=="AMG0-1"]) %>% 
  ungroup() %>%   
  group_by(time, induction) %>% 
  mutate(diff_mean_AbsPro = mean_AbsPro-mean_AbsPro[strain=="AMG0-1"]) %>% 
  mutate(diff_sd_AbsPro = sqrt((sd_AbsPro)^2+(sd_AbsPro[strain=="AMG0-1"])^2)) %>% 
  ungroup() 
```

Plot data generated for alternative robust and sensitive wavelengths
```{r}

#Try plotting the diff as lines instead
plot_diff_lyc_line2 <-
  tbl_lyc_diff2 %>%
  filter(strain == "AlCG0-1") %>%
  #mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
      x = time,
      y = diff_mean_AbsLyc,
      ymin = diff_mean_AbsLyc - diff_sd_AbsLyc,
      ymax = diff_mean_AbsLyc + diff_sd_AbsLyc,
      color = induction
    )
  ) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsLyc, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_fill_manual(values = c("grey", "darkorange"), name = "Induction") +
  scale_color_manual(values = c("grey", "darkorange"), name = "Induction") +
  #geom_errorbar(aes(fill = induction),width=0.2, position=position_dodge(0.9)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" lycopene absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean AlCG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_lyc_line2

plot_diff_pdeox_line2 <-
  tbl_pro_diff2 %>%
  filter(strain == "APDG0-1") %>%
 # mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro,
    color = induction)) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_color_manual(values = c("grey", "darkgreen"), name = "Induction") +
  scale_fill_manual(values = c("grey", "darkgreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" prodeoxyviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APDG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pdeox_line2


plot_diff_pvio_line2 <-
  tbl_pro_diff2 %>%
  filter(strain == "APVG0-1") %>%
 # mutate(time = as.factor(time)) %>%
  mutate(group_rep = paste0(induction, "_", strain)) %>%
  ggplot(
    aes(
    x = time,
    y = diff_mean_AbsPro,
    ymin = diff_mean_AbsPro-diff_sd_AbsPro,
    ymax = diff_mean_AbsPro+diff_sd_AbsPro,
    color = induction)) +
  geom_ribbon(aes(group = induction,
                  fill = induction),
              alpha = 0.2,
              color = NA) +
  geom_line(aes(group = induction)
  ) +
  geom_point(
    aes(x = time, 
        y = diff_AbsPro, 
        color = induction),
    size = 0.5,
    alpha = 0.4
  ) +
  scale_color_manual(values = c("grey", "darkolivegreen"), name = "Induction") +
  scale_fill_manual(values = c("grey", "darkolivegreen"), name = "Induction") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  coord_fixed() +
  ylab(expression(Delta*" proviolacein absorbance [a.u.]")) +
  xlab("Time [h]") +
  ggtitle(expression("Mean APVG0-1 - AMG0-1 (unadjusted OD)"))

plot_diff_pvio_line2

```