---
title: "Investigate difference in absorbance spectra for colorimetric strains"
author: "Andrea Clausen Lind"
date: "2023/05/09"
output: html_notebook
---

## Load libraries
```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(viridis)
library(hrbrthemes)
filter <- dplyr::filter
```

## Load data 
Load data. Adapted to the data format output by the Mars analysis program. (RUN ONCE)
```{r, eval = FALSE}
#Extract file names and paths
Dir_path <- "./2023-06-01_Absorbance_Fluo/"
File_path_list <- list.files(Dir_path, full.names = T)
File_list <- list.files(Dir_path)

sample_time <- 
  File_list %>% 
  as_tibble() %>% 
  mutate(value = gsub("h.xlsx","",value)) %>% 
  mutate(value = gsub(".*UV_","",value)) %>% 
  rename("time" = "value") %>% 
  pull(time)

tbl_full <- tibble()

# Loop through and load each file one by one
for (i in 1:length(File_list)) {
  
  FileToRead <- File_path_list[i]
  FileName <- File_list[i]
  df_file <- read_xlsx(FileToRead, skip = 11, col_names = FALSE) 
  df_file <- 
    df_file %>%  
    select(-"...1", -"...2") %>%
    distinct() %>%
    t() %>%
    row_to_names(row_number = 1) %>%
    as_tibble() %>%
    select(-"Blank B") %>%
    rename(wavelength_nm = "Wavelength [nm]") %>% 
    pivot_longer(!wavelength_nm, names_to = "sample",values_to = "absorbance") %>% 
    mutate(sample = gsub(" ", "_",sample), sample = factor(sample))
  
  df_file <- 
    df_file %>% 
    mutate(time = sample_time[i])
  
  tbl_full <- bind_rows(tbl_full,df_file)
}

# Add metadata
df_meta =
  read_csv("Metadata.csv") %>% 
  mutate(sample = gsub(" ","_",sample))

tbl_complete <- 
  tbl_full %>% 
  full_join(df_meta,by = "sample")

# Tweak columns to numeric and order relevant columns
tbl_complete <- 
  tbl_complete %>%  
  mutate(time = as.integer(time)) %>% 
  mutate(absorbance = as.numeric(absorbance)) %>% 
  mutate(induction = factor(induction,levels = c("0 uM","1 uM"), ordered = T)) %>% 
  mutate(wavelength_nm = as.double(as.character(wavelength_nm))) %>% 
  mutate(strain = factor(strain,levels = c("AMG0-1","AlCG0-1","APDG0-1","APVG0-1"),ordered = T))
  

head(tbl_complete)

saveRDS(tbl_complete,file = "./rds_data/tbl_complete_spectra.rds")
```
Load processed data instead
```{r}
tbl_complete <- readRDS("./rds_data/tbl_complete_spectra.rds")
```

Add unique ID to each biological replicate
```{r}
tbl_complete <- 
  tbl_complete %>% 
  mutate(sample_ID = paste(strain,time,sep = "_")) %>% 
  relocate(sample_ID,.before = wavelength_nm)
```


Calculate average across replicates and conditions. Calculate difference compared to strain without output. 
```{r}

# Check that u have 3 replicates after grouping
tbl_complete %>% 
  group_by(wavelength_nm, sample_ID) %>% 
  mutate(test = length(replicate)) %>% 
  ungroup()

# Add mean and sd to absorbance based on normalized respectively non-normalized data
tbl_complete <- 
  tbl_complete %>% 
  group_by(sample_ID, wavelength_nm, induction) %>% 
  mutate(mean_abs = mean(absorbance),
         sd_abs = sd(absorbance)) %>% 
  ungroup()

# Subtract based on mean (non-normalized). Samples are independent.
tbl_control <- 
  tbl_complete %>% 
  filter(strain == "AMG0-1") %>% 
  select(strain, time, wavelength_nm, mean_abs, induction) %>% 
  distinct() %>% 
  mutate(mean_control = mean_abs) %>% 
  select(-c(mean_abs,strain))

tbl_cases <- 
  tbl_complete %>% 
  select(strain, replicate, time, wavelength_nm, absorbance, mean_abs, induction)

# Caluclate the difference between the negative control and the output strains
tbl_plot <- 
  tbl_control %>% 
  full_join(tbl_cases) %>% 
  mutate(y_plot = absorbance-mean_control,
         y_plot_mean = mean_abs - mean_control)

head(tbl_plot)
```

Plot results for all timepoints and wavelenghts
```{r,fig.width=10}
head(tbl_plot)


plot_diff <- 
  tbl_plot %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_grid(vars(strain),vars(time)) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Absorbance with mean AMG0-1 subtracted")

plot_diff


plot_diff_6h <- 
  tbl_plot %>%
  filter(time == "6") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Absorbance with mean AMG0-1 subtracted at 6 h")

plot_diff_6h

plot_diff_3h <- 
  tbl_plot %>%
  filter(time == "3") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Absorbance with mean AMG0-1 subtracted at 3 h")

plot_diff_3h

plot_diff_10h <- 
  tbl_plot %>%
  filter(time == "10") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Absorbance with mean AMG0-1 subtracted at 10 h")

plot_diff_10h

plot_diff_24h <- 
  tbl_plot %>%
  filter(time == "24") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Absorbance with mean AMG0-1 subtracted at 24 h")

plot_diff_24h



```

Stitch plots and save data
```{r,eval=F}

ggsave(plot_diff_24h, filename = "Figures/DiffAbsSpectra_24h.pdf",width = 7)

ggsave(plot_diff, filename = "Figures/DiffAbsSpectra_All.pdf",width = 10)
```

Make individual plots to match colors
```{r}

plot_diff_22h_lyc <- 
  tbl_plot %>%
  filter(strain %in% c("AlCG0-1")) %>% 
  filter(time == "24") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkorange"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Lycopene absorbance (mean AlCG0-1 - AMG0-1) at 24 h")

plot_diff_24h_lyc

plot_diff_24h_pdeo <- 
  tbl_plot %>%
  filter(strain %in% c("APDG0-1")) %>% 
  filter(time == "24") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkgreen"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Prodeoxyviolacein absorbance (mean APDG0-1 - AMG0-1) at 24 h")

plot_diff_24h_pdeo

plot_diff_24h_pvio <- 
  tbl_plot %>%
  filter(strain %in% c("APVG0-1")) %>% 
  filter(time == "24") %>%
  ggplot(aes(
    x = wavelength_nm,
    y = y_plot_mean)) +
  geom_line(aes(color = induction)) +
  #geom_ribbon(alpha = 0.5,aes(fill = induction)) +
  facet_wrap(vars(strain)) +
  theme_bw() +
    scale_color_manual(values = c("grey","darkolivegreen"), name = "Induction") +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 45,hjust = 1)) +
  ylab(expression(Delta*" Absorbance [a.u.]")) +
  xlab("Wavelength [nm]") +
  ggtitle("Proviolacein absorbance (mean APVG0-1 - AMG0-1) at 24 h")

plot_diff_24h_pdeo

```
Stitch plots and save data
```{r,eval=F}
plot_diff_24h_color <-
  plot_diff_24h_lyc + plot_diff_24h_pdeo + plot_diff_24h_pvio

ggsave(plot_diff_24h_color, filename = "Figures/DiffAbsSpectra_24h_color.pdf",width = 15)

```

